---
title: "Makeup Shades Variability, Price, and Availability  Around the World"
subtitle: "SDS 220 Final Project"
author: 
- name: Katelyn Diaz
  affiliation: Smith College
- name: Sara Gutierrez
  affiliation: Smith College
- name: Allison Wong
  affiliation: Smith College
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 10
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: hide
    link-citations: true
references: 
- id: data.world
  type: article
  title: "data.world Diversity of Makeup Shades"
  author: 
  - family: Lee
    given: Jason 
  issued:
    year: 2018
  publisher: Data.World
  url: https://data.world/the-pudding/diversity-of-makeup-shades
  
- id: Elle
  type: article
  title: "It's 2018. Why Is It Still A Struggle To Find Foundation For Dark Skin?"
  author: 
  - family: Penrose
    given: Nerisha 
  issued:
    year: 2018
  publisher: Elle
  url: https://www.elle.com/beauty/makeup-skin-care/a22687914/struggle-to-find-dark-skin-foundation-woc-influencers-beauty-blender-tarte/
editor_options: 
  chunk_output_type: console
---


```{r setup, message = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  echo = TRUE
)
library(tidyverse)
library(openintro)
library(plotly)
library(moderndive)
library(kableExtra)
library(broom)
makeupdata <- read.csv("makeupshades4 copy.csv")
                       # FinalProjectPersonalExploration/makeupshades1 copy.csv")
  #"SDS220_MakeupShadesStats/MakeupShades3 copy.csv"") <- if i need to run the data
```

```{r filtering out NA in best_seller}
makeup_shades <- makeupdata %>%
  filter(!is.na(best_seller))
```

### Introduction and Explanatory Data Analysis

***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

  During our data exploration, the question that we were trying to answer was if there is a relationship between price and lightness across different value makeup brands around the world. We did this using the data from `Data.World` and their article [Diversity of Makeup Shades](https://data.world/the-pudding/diversity-of-makeup-shades). After playing around with different variables, we settled on `price` and `lightness`, organized by `best_seller`. More specifically, we all liked the idea of finding a relationship between `price` and `lightness`, because we think price is an important factor in determining the likelihood of someone purchasing a makeup product. Since lightness is a number, we can relate it to the price. We also wanted to use the categorical variable `best_seller`, as it allows us to see shade variability separated into four countries, allowing us to observe the diversity of its products and shades, as well as if there is any correlation with `price` in US dollars. 
  
  When glimpsing the data, we noticed about ~100 `na`'s under `best_seller`. To make our analysis easier, we filtered out `na`'s from the data set. Our categorical variable is `best_seller`, and our numerical variables are `price` (quantitative response variable) and `lightness`(explanatory variable). The population that was sampled were best seller beauty brands in four different countries - US, Nigeria, India, and Japan. After consultation of different lists produced from sources such as [Amazon](https://www.amazon.com/Best-Sellers-Beauty-Foundation-Makeup/zgbs/beauty/11058871), [Popsugar Beauty](https://www.popsugar.com/beauty/Bestselling-Foundations-Sephora-2017-43250396), [Refinery29](https://www.refinery29.com/en-us/skin-care-brands-women-of-color), [Beauty In Lagos](https://www.beautyinlagos.com/2017/02/nigerian-makeup-brands-you-should-know-about.html/), and [other sources](https://data.world/the-pudding/diversity-of-makeup-shades), a list of brands were compiled.
  
  To find the shade color, each brand's website was visited in May of 2018, with the color and `price` being recorded based on the liquid foundation line that had the largest variety of shades. Adobe Photoshop was used to determine the `lightness` value.  
  
  Many brands only offer lighter shades because they claim that darker shades are “more difficult to create”, but this is a disservice to makeup consumers with a darker complexion.  We wanted to see if there could be any relationship between the specific shade and price (i.e. are darker shades more expensive?), as well as if this differs by location.

</div>

***

  
### Linear Regressions


#### Parallel Slopes Model and Analysis


```{r}
makeup_model <- lm(price ~ lightness + best_seller, data = makeupdata)
summary(makeup_model)   %>%
  tidy() %>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

> Through this summary model, we are able to better observe our data in the parallel slopes plot. We extraplolated essential information such as the intercepts and slopes of our regression lines in order to better interpret any trends in our data. We also evaluated the p-values, such as the value of approximately 0.02 in row 1, to help us further understand whether or not this fit our dataset properly. 

```{r, fig.cap="**Figure 1a** Parallel slopes model plot comparing lightness (x-axis) to price in dollars (y-axis). India is acting as the baseline (main y-intercept of comparison). Plot compares price and lightness of foundations across four different countries." }
ggplot(makeup_shades, aes(lightness, price, col = best_seller)) + 
    geom_rect(data=NULL,
             aes(xmin=0,xmax=20,ymin=-Inf,ymax=Inf), 
             fill="#ffdbac", col = "#ffdbac")+
  geom_rect(data=NULL,
            aes(xmin=20,xmax=40,ymin=-Inf,ymax=Inf),
            fill="#f1c27d", col="#f1c27d")+
  geom_rect(data=NULL,
            aes(xmin=40,xmax=60,ymin=-Inf,ymax=Inf),
            fill="#e0ac69", col="#e0ac69")+
  geom_rect(data=NULL,
            aes(xmin=60,xmax=80,ymin=-Inf,ymax=Inf),
            fill="#c68642", col="#c68642")+
  geom_rect(data=NULL,
             aes(xmin=80,xmax=100,ymin=-Inf,ymax=Inf), 
             fill="#8d5524", col="#8d5524")+
  scale_x_continuous(breaks=c(0,20,40,60,80,100))+
 scale_color_manual(values=c("#001eff", #india is first
                             "#4deeea", "#ffe700", "#74ee15", "FFFFFF"))+ # color codes the data
 theme_minimal(base_size=20
               # ,
               # scale_fill_discrete(name = "Country")
            )+ # makes a white background, base size changes font size
  labs(title="Parallel Slopes Model of Makeup Data",
       # subtitle="India as the baseline. Looking for correlation between price and lightness\n of foundations across different countries.",
       y= "Price (in US Dollars)",
       x ="Lightness (Lightest to Darkest)") +
  scale_colour_discrete("Best Seller\n by Country") +
  geom_point() +
   geom_parallel_slopes(se=FALSE)
```


```{r, fig.cap="**Figure 1b**Parallel slopes model graph with white background to better view data  points"}
ggplot(makeup_shades, aes(lightness, price, col = best_seller)) +
  scale_x_continuous(breaks=c(0,20,40,60,80,100))+
  theme_minimal(base_size=20)+ 
  labs(title="Parallel Slopes Model of Makeup Data",
       y= "Price (in US Dollars)",
       x ="Lightness (Lightest to Darkest)") +
  scale_colour_discrete("Best Seller\n by Country") +
  geom_point() +
   geom_parallel_slopes(se=FALSE) 
```


***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

For our parallel slopes model, the baseline, India, was represented by the equation: **price_hat_India = 9.1686 -0.007326x**, in which x equals the "lightness" variable. For every 1 unit increase in lightness, we expect to see, on average, an *decrease* of price by approximately $0.007 (the slope), and  when the lightness is equal to zero (the intercept and lightest possible shade) we expect the price to be, on average, **$9.20**. 


Besides from our baseline, we also investigated data for makeup bestsellers in Japan. This was represented by the equation: **price_hat_Japan = (9.2 + 36.72) -0.007326x**. As we can see, the slope stays constant for all lines, so we can expect to see that for every unit increase in lightness, there is also an *decrease* in price by approximately $0.007 on average. When the lightness is equal to zero, we expect to see the price to be, on average, $(9.2 + 36.72), or **$45.92**, so in spite of this constant slope, Japan starts off at a much higher price. This leads us to believe that from the start, the cost of makeup best sellers in Japan is much more expensive than the cost in India. 


Nigeria on the other hand does not appear to show behavior as drastic as Japan. The regression line for Nigeria is represented by **price_hat_Nigeria = (9.2 + 7.88) -0.007326x**, where for every one unit increase in lightness, we expect the price to decrease by $0.007 on average. We also expect the price to be (9.2 + 7.88), or **$17.08** on average, when the lightness is equal to zero. While this price might not be as expensive as the price in Japan, it is still higher than India. 


Lastly, we analyzed at the data for makeup bestsellers in the U.S. The regression line for the U.S. is represented by the equation **price_hat_US = (9.2 + 15.57) -0.007326x**. We expect the price to decrease, on average, by $0.007 for every one unit increase in lightness, and when the lightness is equal to zero, the price is expected to be $(9.2 + 15.57) or **$24.77** on average. Compared to India, the price of makeup best sellers in the U.S. are still more expensive. 


From this analysis, we can see that the price of makeup best sellers in India is the cheapest at $9.20, compared to the other countries which have much higher prices. Japan being the most expensive at $45.92, then the U.S. at $24.77, and finally Nigeria at $17.08. We can see that there is a negative correlation between price and lightness in all countries (India, Japan, Nigeria, the U.S.) as a result of having a negative slope. 

</div>

***

#### Interaction Model and Analysis

```{r}
makeup_model2 <- lm(price ~ lightness * best_seller, data = makeup_shades)
summarymm <- summary(makeup_model2) %>%
  tidy() %>%
  kbl() %>%
  kable_material(c("striped", "hover"))
summarymm
#   tidy() %>%
#   kable()
```

>This summary model describes what is observed in our interaction model plot. We payed close attention to the p-values, which helped us determine that this is not an ideal plot. For example, the p-values of approximately 0.7 in rows 1 and 2 and 0.9 in row 6 which are above the significance level of 0.05 led us to being unable to reject the null hypothesis. 

```{r, fig.cap="**Figure 2** Interaction model linearity plot comparing lightness (x-axis) to price in dollars (y-axis). India is acting as the baseline (main slope of comparison). Plot compares price and lightness of foundations across four different countries. We did not  choose this model."}
ggplot(makeup_shades, aes(lightness, price, col = best_seller))+
  
   geom_rect(data=NULL,
             aes(xmin=0,xmax=20,ymin=-Inf,ymax=Inf), 
             fill="#ffdbac", col = "#ffdbac")+
  geom_rect(data=NULL,
            aes(xmin=20,xmax=40,ymin=-Inf,ymax=Inf),
            fill="#f1c27d", col="#f1c27d")+
  geom_rect(data=NULL,
            aes(xmin=40,xmax=60,ymin=-Inf,ymax=Inf),
            fill="#e0ac69", col="#e0ac69")+
  geom_rect(data=NULL,
            aes(xmin=60,xmax=80,ymin=-Inf,ymax=Inf),
            fill="#c68642", col="#c68642")+
  geom_rect(data=NULL,
             aes(xmin=80,xmax=100,ymin=-Inf,ymax=Inf), 
             fill="#8d5524", col="#8d5524")+
  geom_point() +
  scale_x_continuous(breaks=c(0,20,40,60,80,100))+
  geom_smooth(method = "lm", se=FALSE) +## lm means line, se = FALSE gets rid of the buffer
 scale_color_manual(values=c("#001eff", #india is first
                             "#4deeea", "#ffe700", "#74ee15", "FFFFFF"))+ # color codes the data
 theme_minimal(base_size=20
               # ,
               # scale_fill_discrete(name = "Country")
            )+ # makes a white background, base size changes font size
  labs(title="Interaction Model of Makeup Data",
       # subtitle="India as the baseline. Looking for correlation between price and lightness\n of foundations across different countries.",
       y= "Price (in US Dollars)",
       x ="Lightness (Lightest to Darkest)") +
  scale_colour_discrete("Best Seller\n by Country") 
  
```



***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

At first glance, our interaction model may look similar to our parallel slopes model, but they are actually different. See Appendix for further explanation. 


The interaction model is not ideal compared to the parallel slopes model, and the reasoning behind this lies in the p-values. In the summary table, each row represents a hypothesis test, and the column furthest to the right shows a p-value for each test that was obtained. We usually use a significance level of 0.05 in order to determine whether the result is significant or not. P-values less than 0.05 imply that there is significance (we can reject the null hypothesis because there is evidence in favor in the alternative hypothesis), and those greater than 0.05 do not have significance (we fail to reject the null). As we can see from the summary table for the parallel slopes model, we've obtained p-values all greater than the significance level for seven out of the eight rows (0.7366, 0.7098, and 0.9995, just to name a few). This means that it is not significant to have several different slopes for our model. On the other hand, our parallel slopes model does exhibit extremely small p-values for four out of five of the rows (examples include 0.02, 0.0026, and 3.79e-14). This tells us that there is significance in having different slopes, which is why our parallel slopes model is more ideal. 

</div>

***




### Constant Variance plot

```{r constant variance plot}
ggplot(data = makeup_model, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
   theme_minimal(base_size=20)+
  xlab("Fitted values") +
  ylab("Residuals")
```

***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Our constant variance plot is not ideal. The fitted values represent the price we expect to see from our linear model using the makeup's lightness while the residuals represent the observed price minus the expected price. Each point is a the expected price and how close it was to the original price in the data. An ideal plot would have residual values within a certain range throughout all fitted values. We would be able to draw a rectangle encompassing all the points. This rectangle would show that across all levels of our expected prices, our residuals stay within a certain range and has constant variance. Ideally, the plot would demonstrate as our expected prices from our model increase, we would expect similar differences in our residuals.

However, from our plot, it is clear that a box would not we suit the points. The points instead fit a more triangular shape with a smaller range of residual values (from -7 to 5) at smaller fitted values around an expected price of 10, but a much larger range (-27 to 23) at larger fitted values around an expected price of 45. Starting at lower fitted values, the range of residuals gradually increases until the highest fitted values at 45. Thus, our constant variance plot demonstrates that as our expected prices from our model increase, we also expect greater differences in our residuals. Our model does not demonstrate constant variance. 

</div>

***

### Residuals 


```{r residuals1}
plot1a <- ggplot(makeup_shades, aes(x = lightness, y = price))+
  geom_point()+
  geom_point(aes(color=best_seller))+
  geom_smooth(method = "lm", se = FALSE) # drawing the regression line
#plot1a
#ggplotly(plot1a)
# Summary statistics
question1a <- lm(price ~ lightness, data = makeup_shades)
ggplot(data = question1a, aes(x = .resid)) +
  geom_histogram() +
  xlab("Residuals")+
   theme_minimal(base_size=20)
```


```{r, qq plot}
qqplot <- lm(price ~ lightness, data = makeup_shades)
ggplot(data = question1a, aes(sample = .resid)) +
  geom_qq() +
  xlab("Residuals") +# curves line up with the two peaks in the histogram
   theme_minimal(base_size=20)
```


***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

The residuals histogram can be used to determine if variance is normally distributed. Based on this histogram, it appears that the histogram is right skewed, meaning that they are not normally distributed. This means that we can't trust the p-values on the summary table. 

</div>

***

### Limitations and Conclusion

***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

The question we set out to answer was whether there is a relationship between price and lightness across different value makeup brands around the world. We found no evidence that lightness is associated with price. Although this was not the original question we sought to answer, there is evidence that different countries have different prices in U.S. dollars for makeup because the Pr(>|t|) values on our model's summary table for each individual countries' intercept was much less than the statistical significant value 0.05. If we wanted to continue with this research, we may want to look at the data for makeup prices in more countries to see if makeup prices differ based on shade in different countries.

Our model did not fit well to the metrics used to determine whether or model accurately represents the data. For linearity, each line in our parallel slopes model losely fits the points because many of the points followed alongside the line, but was either had prices extremely higher than the price for the line or lower than that price. Additionally, our model did not have nearly normal residuals or demonstrate constant variance.

Also, there were many limitations in this data set. The data set included many brands that have `na` for `best_seller` value , and we filtered all of those brands out of our data. Because we filtered out these brands, it may not be appropriate to generalize these results to all brands or even all countries because there were only 4 countries that were best sellers in the data set. Makeup prices may not vary greatly as lightness increases in a country in our data set, but in another country not represented by our data, it still could vary greatly. Similarly, a makeup brand in our data may not accurately show how prices for different shades of lightness vary in a different brand.

</div>

***

...

### Appendix

***
<style>
div.blue { background-color:#f2f1f0; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

*Further explanations for the Interaction Model:* 

For our interaction model, the baseline, India, was represented by the equation: **price_hat_India = 4.1129345 + 0.0607310x**, in which x equals the "lightness" variable. For every 1 unit increase in lightness, we expect to see, on average, an increase of price by $0.06 (the slope), and  when the lightness is equal to zero (the intercept and lightest possible shade) we expect the price to be, on average, $4.11. Because this is the interaction model, we can expect our intercepts and slopes to change, depending on the country we are looking at. 

Besides from our baseline, we also looked at data for makeup bestsellers in Japan. This was represented by the equation: **price_hat_Japan = (4.11 + 36.55) + (0.0607310 - 0.0001561)x**. On average, we can expect to see that for every unit increase in lightness, there is an increase in price by $(0.0607310 - 0.0001561), or approximately $0.06. When the lightness is equal to zero, we expect to see the price to be, on average, $(4.11 + 36.55), or $40.66. Even though the slope of the baseline and the slope of **price_hat_Japan** are very similar, Japan is already at a much higher price.


Nigeria on the other hand does not appear to show behavior as extreme as Japan. The regression line for Nigeria is represented by **price_hat_Nigeria = (4.11 + 12.64) + (0.0607310 - 0.0626467)x**, where for every one unit increase in lightness, we expect the price to *decrease* by $(0.0607310 - 0.0626467), or $0.0019 on average. The slope of **price_hat_Nigeria** is roughly the same as the baseline, since that value isn't even equal to one cent. When the lightness is equal to zero, we expect the price to be (4.11 + 12.64), or $16.75 on average. While this price might not be as expensive as the price in Japan, it is still higher than the baseline. 


Lastly, we looked at the data for makeup bestsellers in the U.S. The regression line for the U.S. is represented by **price_hat_US = (4.11 + 21.73) + (0.0607310 - 0.0851736)x**. We expect the price to decrease, on average, by $(0.0607310 - 0.0851736) or approximately $0.03 for every one unit increase in lightness. When the lightness is equal to zero, the price is expected to be $(4.11 + 21.73) or $25.84 on average. Compared to the baseline, the price of makeup best sellers in the U.S. are still more expensive to those in India. 


In general, we can see that the price of makeup best sellers in India is the cheapest at $4.11, compared to the other countries which have much higher prices. Japan being the most expensive at $40.66, then the U.S. at $25.84, and finally Nigeria at $16.75. From the slopes, we can see that in Japan and India, there is a positive correlation between price and lightness, and then in Nigeria and the U.S., a negative correlation between the two. 

</div>

***


### References

